---
title: "IR-SPG: Example"
output: html_vignette
vignette: >
  %\VignetteIndexEntry{multifractal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## The Task

Find the nearest correlation matrix subject to an equality constrain. This problem was described in Schadner and Traut (2022) at the estimation of forward-looking stock correlation matrices.

## Solving the Problem

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The test sample covers implied volatilities, weights, Pearson's correlations and pre-computed values for the cross-section and index of the S&P 100 as observed on January 31, 1996. 

#### 1. Defining the Functions

```{r}

## Objective Function
# i.e., the function to be minimized
objfn <- function(x){
  x <- matrix(x, n)
  norm(J * x %*% t(x)  - Ahat, "F")^2
}

# gradient of objective function
objfn_g <- function(x){
  x <- matrix(x, n)
  c(4*(J * x %*% t(x) - Ahat) %*% x)
}

## Equality Constraint
# i.e., the C(x) function within Gomes-Ruggiero et al. (2009)
eqfn <- function(x){
  x <- matrix(x, n)
  c(t(w) %*% sig %*% (x %*% t(x) * J) %*% sig %*% w - vmb)
}

# gradient of equality constraint
eqfn_g <- function(x){
  x <- matrix(x, n)
  c(- 2* sig %*% (J * w %*% t(w)) %*% sig %*% x)
}

## Projection Function
# i.e., the projection on the tangible set
# see Gomes-Ruggiero et al. (2009) for greater details
projfn <- function(y){
  Peq <- function(Y){
    XX <- Y%*%t(Y)
    aa <- c(t(b) %*% (B %*% XX %*% B * J) %*% b)
    bb <- 2*c(t(b)%*%((XX %*% B)*J) %*% b)
    cc <- c(t(b) %*% (XX * J) %*% b - vex)
    dd <- max(0, bb^2 - 4*aa*cc)
    lam <- (-bb + sqrt(dd))/(2*aa)
    return((I + lam*B) %*% Y)
  }
  
  y <- matrix(y, n)
  for(kk in 1:100){
    y <- Peq(y)
    rs <- rowSums(y^2)
    if(max(rs) > 1.02){
      ind <- sqrt(rs)
      ind[ind < 1] <- 1
      ind <- ind + (ind - 1) * 0.1
      y <- y/ind
    } else {
      break
    }
    if(abs(eqfn(y)) < vtol) break
  }
  return(c(y))
}
```

#### 2. Load the Data and Run the Optimization

```{r}
library(irspg)

## loading and assigning the sample data
data(sp100) 
tmp <- sapply(ls(), function(z) assign(z, get(z))) 

## Run the optimization
mdl <- irspg(x0, objfn, objfn_g, projfn, eqfn, eqfn_g)


## Results

# (preview) the best set of parameters found
head(mdl$X) 

# final value of objective function
mdl$objfn

# starting value of objective function
mdl$opti_at_first

# accuracy of meeting the equality constraint
mdl$eq_error

# computational time
mdl$ctime


```
## References

Schadner, W & Traut, J. (2022). "Estimating Forward-Looking Stock Correlations from Risk Factors", *Mathematics* 10(10), 1649. <https://doi.org/10.3390/math10101649>.

